An Automatic Verification Framework for Software Configurations

### 2016-05-04 OSDI/arch.tex
config: new abs
config: updated

\section{\app Overview}

\begin{figure*}[tbp] \centering
\includegraphics[width=0.68\textwidth]{figs/overview}
\caption{{\footnotesize \app's workflow. The green boxes represent configuration files,
  including both correct general configuration files and users' input
  configuration files. The purple boxes are the components within \app.
  The yellow boxes are results generated by \app's components.}}
\label{fig-overview}
\end{figure*}

Figure~\ref{fig-overview} describes an overview of our system. We start
with the assumption that we are given a number of correct configuration
files belonging to the same category, such as MySQL or Apache. 
Such files follow similar patterns, which we exploit
in a learning algorithm to build rules that
describe a language model for the files. Since the
``language'' of configuration types is untyped and unstructured, we
first parse the files and translate them into a more structured,
intermediary representation.
When running type inference on a configuration file, 
the type of a variable cannot always be fully determined from 
a single value.
We address this problem 
by introducing so called {\emph{probabilistic types}}.
Rather than giving a variable a single type, we assign several types with their probability distributions. 
We can then use these more structured files
as a training set to learn the rules. The learning algorithm
is template-based to be easily extensible. We provide an initial set of templates and the
learner learns some concrete instances from the training set. These
rules are used for detecting errors violating the learned constraints
in the files given by the user.

As an 
illustration of a simple rule that we can learn, consider a template
 $X_1 \le X_2$, where $X_1$ and $X_2$ are
integer variables. The learner might derive the rule stating that
$\texttt{mysql.max\_persistent} \le \texttt{max\_connections}$. There is a classification and taxonomy of configuration errors in the 
existing work on automated configuration troubleshooting~\cite{yin11anempirical, configdataset}. We provide templates for every class that \app can handle: we consider integer constraints, ordering
constraints, typing constraints, and constraints about correlated entries (such as ``if $X$ is present, $Y$ has to appear as well''). 
Unfortunately, we cannot handle the class of errors that rely on the analysis of the whole operating system.
Our language-based approach can only learn on sets of text files, not the system environment.

\com{
A rule R is added to the set of all rules,
  if $\exists$ learning file f s.t. R(f) is non vacuously true
That rule R is then removed from the set of all rules,
  if $\exists$ learning file f st R(f) is false.

This is accomplished in two passes.
First we collect all possible rules for every file.
Then we merge the all the rules to create our final set.
This conviently gives rise to an embarresingly parallel situation, which Haskell allows us to easily take advantage of by
  using the parallel mapping library parmap.

\begin{lstlisting}
  potentialRules = parmap findAllRules learningSet.
  finalRules = foldl1 mergeRules potenialRules
\begin{lstlisting}

Each rule is of the Attribute typeclass, which means a rule must support the following operations:
\begin{lstlisting}
class Attribute r where
  learn :: ConfigFile Common -> [r]
  merge :: [r] -> [r] -> [r] 
  check :: [r] -> ConfigFile Common -> Error
\end{lstlisting}

}

>>>>

\section{The \app Framework Overview}

\begin{figure*}[tbp] \centering
\includegraphics[width=0.78\textwidth]{figs/overview}
\caption{\app's workflow. The green boxes represent configuration files,
  including both correct general configuration files and users' input
  configuration files. The purple boxes are the components within \app.
  The yellow boxes are results generated by \app's components.}
\label{fig-overview}
\end{figure*}

We propose \app, an automatic verification framework,
which can solve configuration errors, \eg, ordering errors
and fine-grained value correlation, that previous work cannot tackle.
As depicted in Figure~\ref{fig-overview}, \app has three main steps:
translation, learning and checking. In this section, we briefly
describe how does each step work.

\para{Initial state.}
We start
with the assumption that we are given a number of (not necessarily) 
correct configuration files belonging to the same system, 
such as MySQL or Apache. 
Such files follow similar patterns, which we exploit
in a learning algorithm to build rules that
describe a language model for the files.

\para{Translator.}
The translator component first parses the input sample 
dataset (containing both configuration files and system environment
information), and then transforms them to a more structured
and typed intermediate representation.
When we run type inference on a configuration file, 
the type of a variable cannot always be fully determined from 
a single value.
We address this problem 
by introducing so called {\em probabilistic types}.
Rather than giving a variable a single type, 
we assign several types with their probability distributions. 
We can later use these more structured files
as a training set to learn the rules.

\para{Learning.}
The learning algorithm is template-based to be easily extensible. 
We provide an initial set of templates and the
learner learns some concrete instances from the training set. These
rules are used for detecting errors violating the learned constraints
in the files given by the user.

As an illustration of a simple rule that we can learn, consider a template
 $X_1 \le X_2$, where $X_1$ and $X_2$ are
integer variables. The learner might derive the rule stating that
$\texttt{mysql.max\_persistent} \le \texttt{max\_connections}$. There is a classification and taxonomy of configuration errors in the 
existing work on automated configuration troubleshooting~\cite{yin11anempirical, configdataset}. We provide templates for every class that \app can handle: we consider integer constraints, ordering
constraints, typing constraints, and constraints about correlated entries (such as ``if $X$ is present, $Y$ has to appear as well'').

### 2016-05-04 OSDI/intro.tex
config: updated

While many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{attariyan10automating,
su07autobash, whitaker04configuration}, 
offering automatic verification to configuration files -- like what we
did for verifying regular programs -- is still highly
desirable~\cite{wang04automatic, zhang14encore}.
Nevertheless, the main obstacles are:
1) entries in configuration files are untyped entries, 2) there
is no explicit structure policy for the entries in configuration
files, and 3) there are surprisingly few rules specifying the
entries' constraints.

In order to avoid these obstacles,
researchers propose statistical analysis and learning based approaches%
~\cite{wang04automatic, zhang14encore, yuan11context}. 
These efforts build checking policies by learning a sample 
data set, rather than explicitly specifying entries' types or rules.
In particular, for each entry in a certain configuration file, 
if it deviates from the common values used in a large collection
of configurations (\ie, the training set), it is typically
suspected as a potential configuration error.
However, because these learning 
efforts {\em either} are limited to simplistic 
configuration errors (\eg, type errors and syntax errors), 
{\em or} heavily rely on template-based inference~\cite{zhang14encore}, 
many sophisticated configuration errors, 
\eg, hard to be templated in practice, cannot be detected.
For example, if {\tt extension = mysql.so} appears 
before {\tt extension = recode.so} in PHP configuration file,  
it would lead to a crash error, since the correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical};
however, such an error cannot be detected by existing
learning efforts, since it is hard to build a template for this
error~\cite{xu15systems}.

@@

Based on the above insight, 
this paper presents \app, the first automatic verification framework
for general software configurations.
In particular, \app's workflow to verifying a given configuration file
could be looked as a three-step methodology.
First, \app analyzes a dataset containing sample configuration files,
thus generating a well-structured and probabilistically-typed 
intermediate representation.
Second, \app derives rules and constraints by analyzing
the intermediate representations, thus building a language model.
Finally, \app uses the resulting language model
to verify the given configuration file and detect potential errors.
Compared with previous efforts,
\app does not necessarily rely on users' templates, 
and is capable of detecting more tricky configuration errors that
cannot be identified by existing work (details in $\S$\ref{sec-motiv}).

Building such an automatic verification framework for
configuration files, nevertheless, requires addressing several challenges. 
First, before formulating the language model, 
we need to infer each entry's type in a given configuration file;
however, the type of a variable cannot always be fully determined 
from a single value. 
For example, an entry {\tt foo = MAX\_SIZE} is most likely
an integer rather than string. We address this problem by introducing 
so called {\emph{probabilistic types}}.
Rather than assigning only one variable to a single type,
like previous work did~\cite{zhang14encore}, 
we assign several types with their probability distributions. 
The entry in the above example should be assigned 
a probabilistic type like 
{\tt \{foo, MAX\_SIZE, [(Int, 90\%), (String, 10\%)]\}}.
\ennan{I am not quite sure probabilistic types are obviously
better than deterministic types.}

Second, without template, how to learn rules and constraints present
a difficult. \ennan{Here, we need to add a describe to illustrate
how we learn rules without templates.}

@@

Our \app prototype still has a few limitations,
\eg, cannot handle errors occurred in execution time.
Nevertheless, we believe \app may suggest a practical path
toward automatic and language-based configuration verification.

@@

To summarize, this tool paper makes the following contributions:

@@

\item We propose the first automatic configuration verification
framework, \app, that can learn a language model from an example set of 
correct configuration files, and we use the model to verify 
new configuration files.

\item \app uses probabilistic types to assign a confidence 
distribution over a set of types to each entry, 
while generating the intermediate representation.

\item \app applies existing machine learning algorithms to 
enable powerful rule and constraint inference without the assistance 
from any pre-defined templates.

>>>>

While many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{attariyan10automating,
su07autobash, whitaker04configuration}, 
offering automatic verification to configuration files -- like what we
did for verifying regular programs -- is still highly
desirable~\cite{wang04automatic, zhang14encore, xu15systems}.
Nevertheless, the main obstacles to the configuration verification are:
1) entries in configuration files are untyped entries, 2) there
is no explicit structure policy for the entries in configuration
files, and 3) there are surprisingly few rules specifying the
entries' constraints.

In order to overcome these obstacles,
researchers propose statistical analysis and learning based approaches%
~\cite{wang04automatic, zhang14encore, yuan11context}. 
These efforts build checking policies by learning a sample 
data set, rather than explicitly specifying entries' types or rules.
In particular, for each entry in a certain configuration file, 
if it deviates from the common values used in a large collection
of configurations (\ie, the training data set), it is typically
suspected as a potential configuration error.
However, because these learning 
efforts {\em either} are limited to simplistic 
configuration errors (\eg, type errors and syntax errors), 
{\em or} heavily rely on template-based inference~\cite{zhang14encore}, 
many sophisticated configuration errors, 
\eg, hard to be templated in practice, cannot be detected.
For example, if {\tt extension = mysql.so} appears 
before {\tt extension = recode.so} in PHP configuration file,  
it would lead to a crash error, since the correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical};
however, such an error cannot be detected by existing
learning efforts, since it is hard to build a template for this
error~\cite{xu15systems}.

@@

Based on the above argument, 
this paper presents \app, the first automatic verification framework
for general software configurations.
In particular, \app's workflow to verifying a given configuration file
could be looked as a three-step methodology.
First, \app analyzes a dataset containing sample configuration files,
thus generating a well-structured and probabilistically-typed 
intermediate representation.
Second, \app derives rules and constraints by analyzing
the intermediate representations, thus building a language model.
Finally, \app uses the resulting language model
to verify the given configuration file and detect potential errors.
Compared with previous efforts,
\app does not necessarily rely on users' templates, 
and is capable of detecting more tricky configuration errors that
cannot be identified by existing work (details in $\S$\ref{sec-motiv}).

Building such an automatic verification framework for
configuration files, nevertheless, requires addressing several challenges. 
First, in order to formulate a correct language model, 
we need to infer each entry's type in a given configuration file;
however, the type of a variable cannot always be fully determined 
from a single value. 
For example, an entry {\tt foo = MAX\_SIZE} is most likely
an integer rather than string; however, existing type inference 
work would report this is an error, because foo should be assigned
an integer~\cite{zhang14encore}. We address this problem by introducing 
so called {\emph{probabilistic types}}.
Rather than assigning only one variable to a single type, 
we assign several types with their probability distributions. 
The entry in the above example might be assigned 
a probabilistic type like 
{\tt \{foo, MAX\_SIZE, [(Int, 60\%), (String, 40\%)]\}}.
With such probabilistic types in hand,
we can generate a more accurate language model,
thus significantly improving our checking capability.

Second, without template, how to learn rules and constraints present
a difficult. \ennan{Here, we need to add one-paragraph description 
to illustrate how we learn rules without templates.}

@@

Our \app prototype still has a few limitations,
\eg, cannot handle errors occurred in execution time.
Nevertheless, we believe \app may suggest a practical path
toward automatic and language-based configuration verification.
To summarize, this tool paper makes the following contributions:

@@

@@

\item We propose the first automatic configuration verification
framework, \app, that can learn a language model from an example set of 
correct configuration files, and then uses this language model to verify 
interested configuration files.

\item \app proposes probabilistic types to assign a confidence 
distribution over a set of types to each entry, 
while generating the intermediate representation.

\item \app employs a collection of machine learning algorithms to 
enable powerful rule and constraint inference without the assistance 
from any pre-defined templates.

### 2016-05-04 OSDI/limit.tex
config: updated

\ennan{Of course, our tool has many limitations. For example, we cannot
detect the misconfiguration correlation across multiple individual
systems unwilling to share configuration information with each other.}

@@

\para{Legal misconfigurations.}
While \app can check diverse configuration errors without
human participations, all the misconfiguration detection techniques,
including \app, cannot handle some types of configuration errors
resulting from events occurred during system runtime.
Such configuration errors are referred to as {\em legal misconfigurations}%
~\cite{yin11anempirical}. As shown in Yin {\em et al.}%
~\cite{yin11anempirical}, many parameter misconfigurations have 
perfectly legal parameters but do not deliver the functionality intended
by users. These cases are more difficult to detect by
automatic checkers and may require more user training or
better configuration design.
A potential solution is to combine existing misconfiguration diagnosis
tools, \eg, X-ray~\cite{}, with \app in order to enhance the 
misconfiguration checking capability.

\para{}

>>>>

@@

\para{Legal misconfigurations.}
While \app can check diverse configuration errors without
human participations, all the misconfiguration detection techniques,
including \app, cannot handle some types of configuration errors
resulting from events occurred during system runtime.
Such configuration errors are referred to as {\em legal misconfigurations}%
~\cite{yin11anempirical}. As shown in Yin {\em et al.}%
~\cite{yin11anempirical}, many parameter misconfigurations have 
perfectly legal parameters but do not deliver the functionality intended
by users. These cases are more difficult to detect by
automatic checkers and may require more user training or
better configuration design.
A potential solution is to combine existing misconfiguration diagnosis
tools, \eg, X-ray~\cite{attariyan12x-ray}, 
with \app in order to enhance the 
misconfiguration checking capability.

\para{Misconfiguration across software components.}
As exposed by Yin {\em et al.}~\cite{yin11anempirical},
cross-software configuration correlation problems also account
for a considerable number of misconfiguration cases.
For example, in a LAMP-based Web server, one entry in 
PHP configuration file, {\tt mysql.max\_persistent = 400}
may make users encounter a ``too many connections'' error,
because a correlated entry in the underlying MySQL's configuration
file assigns {\tt max\_connection} to 300, which is less
than the MySQL connection numbers in PHP's configuration file (\ie, 400).
It is quite difficult to detect such a type of tricky error
through leaning approaches, because not only users or engineers 
are not aware of the hidden interactions~\cite{xu15systems},
but also it is hard to obtain a global knowledge to the entire
configurations.
One possible solution to this problem might be to introduce
some cryptographic protocol, \eg, private set
intersection~\cite{kissner05privacy}, to privately extract the
overlapping entries, \eg, {\tt mysql.max\_connection} in the 
above MySQL and PHP case, for double-checking.

\para{Network configuration verification.}
\app mainly focuses on software configurations, \eg, MySQL and Apache,
so that our approach is limited to support network configuration
verification. This is because network configurations have quite
different representations, format and rules from software configurations,
since network configurations are typically written in 
more domain-specific policy languages.
In fact, many network verification tools, 
\eg, NoD~\cite{lopes15checking} and 
Dobrescu {\em et al.}~\cite{dobrescu14software},
have been proposed to check whether network configurations
meet their specifications.

### 2016-05-04 OSDI/motiv.tex
config: updated

In this section, we present several tricky configuration errors
extracted from real-world examples. 
Existing efforts, \eg, learning-based checking
efforts~\cite{zhang14encore, wang04automatic}, cannot detect
these sophisticated misconfiguration problems.

\para{Example~1: Ordering errors.} 
The first example configuration error is extracted from 
Yin {\em et al.}~\cite{yin11anempirical}.
When a user configure PHP to run with the
Apache HTTP Server, this user writes, among others, the following lines
in the configuration file.

@@

Such a configuration file will cause the Apache server to 
fail to start due to a segmentation fault error~\cite{yin11anempirical}. 
This is because when using PHP in Apache, the extension {\tt mysql.so} 
depends on {\tt recode.so}, and the relative ordering of two of them 
is crucial. \app is able to inform the user that {\tt recode.so} 
should appear before {\tt mysql.so}, and reports the error.

@@

\ennan{Example~4: This should be the singular value error.}

\para{Other errors.}
\app can also deal with the configuration errors that can be detected by
existing work, such as EnCore~\cite{zhang14encore}.
For example, both \app and EnCore can handle system execution related
misconfiguration problems.
Furthermore, \app is also able to detect type errors and syntax errors
in configuration files.

>>>>

In this section, we present several {\em tricky} configuration errors
extracted from real-world examples. 
The reason we call them tricky is
most of existing efforts, \eg, learning-based checking
efforts~\cite{zhang14encore, wang04automatic}, cannot detect
these configuration errors.
On the other hand, we also investigated 
many misconfiguration problems in practice were caused by 
these tricky configuration errors, \eg, reported by Stack Overflow.

\para{Example~1: Ordering errors.} 
The first example configuration error is first reported by 
Yin {\em et al.}~\cite{yin11anempirical}.
For instance, when a user configures PHP to run with the
Apache HTTP Server, this user may write, among others, 
the following lines in the configuration file.

@@

Such a configuration file will cause the Apache server to 
fail to start due to a segmentation fault error~\cite{yin11anempirical}. 
This is because when using PHP in Apache, the extension {\tt mysql.so} 
depends on {\tt recode.so}, and the relative ordering of two of them 
is crucial. 
We call the above example misconfiguration situation 
as {\em ordering errors}.
Yin {\em et al.} find that ordering errors widely exist in
many system configurations, \eg, PHP and MySQL,
and typically lead to multiple system crash events;
however, there is no existing effort that can effectively solve 
or detect this problem~\cite{zhang14encore, xu15systems, xu13do}.

If this user is using \app, she can avoid such a configuration error.
In particular, \app is able to inform the user that {\tt recode.so} 
should appear before {\tt mysql.so}, and reports the error (as shown
below).

@@

\ennan{Example~4: This should be the singular value error.
We need an example from Stack Overflow to illustrate the 
real-world problem. I guess the buffer size example is fine.}

\para{Other errors.}
\app can also deal with the configuration errors that can be detected by
existing work, such as EnCore~\cite{zhang14encore}.
For example, both \app and EnCore can handle system execution related
misconfiguration problems.
Furthermore, \app is also able to detect type errors and syntax errors
in configuration files.
\ennan{We may need one example here.}
